---
title: "index"
author: "Robert Esposito"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,"load-libraries"}
library(tidyverse)
library(languageR)
```

```{r,"view-languageR"}
data(package="languageR")
?durationsGe
durationsGe ## df is tidy
```

```{r,"descriptive-statistics"}
## five-number summary
summary(durationsGe$DurationOfPrefix)
summary(durationsGe$SpeechRate)

## standard deviation
sd(durationsGe$DurationOfPrefix)
sd(durationsGe$SpeechRate)

## convert to z scores
z_scores <- durationsGe %>%
  mutate(
    z_DurationOfPrefix = (DurationOfPrefix-mean(DurationOfPrefix)/sd(DurationOfPrefix)),
    z_SpeechRate = (SpeechRate-mean(SpeechRate)/sd(SpeechRate)))
z_scores

## plot z scores
## I don't think this is particularly useful, especially on the same graph, since I'm looking at two continuous variables and I'm not looking at a group
z_scores %>% 
  ggplot() +
  geom_histogram(aes(z_DurationOfPrefix),fill="red",bins=500) +
  geom_histogram(aes(z_SpeechRate),fill="blue",bins=100) +
  labs(x=NULL)

## the same plots, but more readable
z_scores %>% 
  ggplot() +
  geom_histogram(aes(z_DurationOfPrefix),fill="red")

z_scores %>% 
  ggplot() +
  geom_histogram(aes(z_SpeechRate),fill="blue",bins=100)
```

```{r,"bivariate-regression"}
## fitting the linear model
summary(lm(DurationOfPrefix ~ SpeechRate,data=durationsGe))

## correlation
## dor = duration of prefix
## sr = speech rate
corr_dor_by_sr <- cor(durationsGe$DurationOfPrefix, durationsGe$SpeechRate)
corr_dor_by_sr
```

```{r,"plot-with-regression-line"}
plot_duration_by_sr <- durationsGe %>% ggplot() +
  aes(x=SpeechRate,y=DurationOfPrefix) +
  geom_smooth(method="lm") +
  geom_point()
plot_duration_by_sr
```

I wanted to investigate the relationship between speech rate and the duration of the Dutch prefix <ge->. To do this, I fit a linear model to the two variables, running the duration of the prefix as a function of speech rate. The model displayed a negative relationship: as speech rate increases, the duration of the prefix decreases. The correlation was not particularly strong at `r corr_dor_by_sr`. 

However, the summary statistics for the linear model show that there is a significant relationship (p<.001) between the two variables. It is surprising to me that the correlation can be so low, but the p-value for the linear model can be significant at such a confidence interval. I would be interested to know if certain affixes or parts of speech are more or less resistant to change based on speech rate.